---
title: "Introduction"
output: html_notebook
---

We import the data from the [official page of the dataset](https://www.uni-goettingen.de/en/551625.html).

## Example 1.1: Munich Rent Index

```{r}
munich_rent_url <- "https://www.uni-goettingen.de/de/document/download/64c29c1b1fccb142cfa8f29a942a9e05.raw/rent99.raw"

munich_rent_index <- read.table(
  url(munich_rent_url),
  header = 1,
  colClasses = c(
    "numeric", "numeric", "numeric",
    "numeric", "factor", "factor",
    "factor", "factor", "factor"
  )
)

# Convert to logical
munich_rent_index$bath <- munich_rent_index$bath == 1
munich_rent_index$kitchen <- munich_rent_index$kitchen == 1
munich_rent_index$cheating <- munich_rent_index$cheating == 1
summary(munich_rent_index)
```

For future use, this code is going to be available in an R script in [`import_data/munich_rent_index.R`](import_data/munich_rent_index.R).

# Example 2.2: Malnutrition in Zambia

```{r}
zambia_url <- "https://www.uni-goettingen.de/de/document/download/d90a2d7b26c4504ab6630cf36cbae2fa.raw/zambia_height92.raw"

malnutrition_zambia <- read.table(
  url(zambia_url),
  header = 1,
  colClasses = c(
    "numeric", "factor", "numeric",
    "integer", "numeric", "numeric",
    "numeric", "factor", "factor",
    "factor", "factor", "factor"
  )
)

# Convert to logical
malnutrition_zambia$m_work <- malnutrition_zambia$m_work == 1
summary(malnutrition_zambia)
```

Code for import in [`import_data/malnutrition_zambia.R`](import_data/malnutrition_zambia.R).

## Example 1.3: Patent Opposition

```{r}
patent_url <- "https://www.uni-goettingen.de/de/document/download/66eb4eb0bc0e8f6acf1d02ddf683f077.raw/patentdata.raw"

patent_opposition <- read.table(
  url(patent_url),
  header = 1,
  colClasses = c(
    "factor", "factor", "factor",
    "factor", "factor", "integer",
    "integer", "integer", "integer"
  )
)
# Convert to logical
patent_opposition$biopharm <- patent_opposition$biopharm == 1
patent_opposition$ustwin <- patent_opposition$ustwin == 1
patent_opposition$patus <- patent_opposition$patus == 1
patent_opposition$patgsgr <- patent_opposition$patgsgr == 1
summary(patent_opposition)
```

R code in [`import_data/patent_opposition.R`](import_data/patent_opposition.R). 

## Example 1.4: Forest Health Status

```{r}
forest_url <- "https://www.uni-goettingen.de/de/document/download/f5ef58e05aff8b6546dcf993aa73a480.raw/beach.raw"

forest_health_status <- read.table(
  url(forest_url),
  header = 1,
  colClasses = c(
    "factor", "numeric", "factor",
    "numeric", "numeric", "numeric",
    "numeric", "numeric", "numeric",
    "numeric", "numeric", "factor",
    "factor", "factor", "factor",
    "factor"
  ),
  na.strings = "."
)
summary(forest_health_status)
```

R code in [`import_data/forest_health_status.R`](import_data/forest_health_status.R).

## Example 1.5: Munich Rent Index - Univariate Distributions

```{r, fig.width=10, fig.height=8, fig.align='center'}
par(mfrow = c(2, 2))

ylab <- 'estimated density'
hist(munich_rent_index$rent, freq=FALSE, xlab = 'net rent in Euro', ylab = ylab)
lines(density(munich_rent_index$rent), col = "red", lwd = 2)

hist(munich_rent_index$rentsqm, freq=FALSE, xlab = 'net rent per sqm in Euro', ylab = ylab)
lines(density(munich_rent_index$rentsqm), col = "red", lwd = 2)

hist(munich_rent_index$area, freq=FALSE, xlab = 'area ub sqm', ylab = ylab)
lines(density(munich_rent_index$area), col = "red", lwd = 2)

hist(munich_rent_index$yearc, freq=FALSE, xlab = 'year of construction', ylab = ylab)
lines(density(munich_rent_index$yearc), col = "red", lwd = 2)
```

## Example 1.6: Malnutrition in Zambia - Univariate Distributions

```{r, fig.width=10, fig.height=12, fig.align='center'}
par(mfrow = c(3, 2))

ylab <- 'estimated density'
hist(malnutrition_zambia$zscore, freq=FALSE, xlab = 'child\'s Z-score', ylab = ylab)
lines(density(malnutrition_zambia$zscore), col = "red", lwd = 2)

hist(malnutrition_zambia$c_breastf, freq=FALSE, xlab = 'duration of breast feeding in month', ylab = ylab)
lines(density(malnutrition_zambia$c_breastf), col = "red", lwd = 2)

hist(malnutrition_zambia$c_age, freq=FALSE, xlab = 'child\'s age in months', ylab = ylab)
lines(density(malnutrition_zambia$c_age), col = "red", lwd = 2)

hist(malnutrition_zambia$m_agebirth, freq=FALSE, xlab = 'mother\'s age at birth', ylab = ylab)
lines(density(malnutrition_zambia$m_agebirth), col = "red", lwd = 2)

hist(malnutrition_zambia$m_height, freq=FALSE, xlab = 'mother\'s height in cm', ylab = ylab)
lines(density(malnutrition_zambia$m_height), col = "red", lwd = 2)

hist(malnutrition_zambia$m_bmi, freq=FALSE, xlab = 'mother\'s BMI', ylab = ylab)
lines(density(malnutrition_zambia$m_bmi), col = "red", lwd = 2)
```

## Example 1.7: Munich Rent Index - Scatter Plots

```{r, fig.width=10, fig.height=8, fig.align='center'}
par(mfrow = c(2, 2))

plot(
  x = munich_rent_index$area,
  y = munich_rent_index$rent,
  xlab = "area in sqm",
  ylab = "net rent in Euro"
)

plot(
  x = munich_rent_index$area,
  y = munich_rent_index$rentsqm,
  xlab = "area in sqm",
  ylab = "net rent per sqm in Euro"
)

plot(
  x = munich_rent_index$yearc,
  y = munich_rent_index$rent,
  xlab = "year of construction",
  ylab = "net rent in Euro"
)

plot(
  x = munich_rent_index$yearc,
  y = munich_rent_index$rentsqm,
  xlab = "year of construction",
  ylab = "net rent per sqm in Euro"
)
```

## Example 1.8: Munich Rent Index - Clusttered Scatter Plots

```{r, fig.width=10, fig.height=8, fig.align='center'}
par(mfrow = c(2, 2))

boxplot(
  rent ~ area,
  data = munich_rent_index,
  xlab = "area in sqm",
  ylab = "net rent in Euro",
  xaxt = "n"
)

boxplot(
  rentsqm ~ area,
  data = munich_rent_index,
  xlab = "area in sqm",
  ylab = "net rent per sqm in Euro",
  xaxt = "n"
)

boxplot(
  rent ~ yearc,
  data = munich_rent_index,
  xlab = "year of construction",
  ylab = "net rent in Euro",
  xaxt = "n"
)

boxplot(
  rentsqm ~ yearc,
  data = munich_rent_index,
  xlab = "year of construction",
  ylab = "net rent per sqm in Euro",
  xaxt = "n"
)
```

## Example 1.9: Munich Rent Index - Categorical Explanatory Variables

```{r, fig.width=10, fig.height=4, fig.align='center'}
par(mfrow = c(1, 2))

boxplot(
  rentsqm ~ location,
  data = munich_rent_index,
  xlab = '',
  ylab = 'net rennt per sqm',
  names = c('average', 'good', 'top')
)

plot(density(subset(munich_rent_index, location == 1)$rentsqm), xlab = 'net rent per sqm', ylab = 'estimated density', lwd = 2)
lines(density(subset(munich_rent_index, location == 2)$rentsqm), col = "green", lwd= 2)
lines(density(subset(munich_rent_index, location == 3)$rentsqm), col = "red", lwd= 2)

legend(
  "topright",
  legend = c('average', 'good', 'top'),
  col = c('black', 'green', 'red'),
  lwd = 2
)
```

## Example 1.10: Malnutrition in Zambia - Graphical Association Analysis

```{r, fig.width=10, fig.height=4, fig.align='center'}
par(mfrow = c(1, 2))

plot(
  x = malnutrition_zambia$c_age,
  y = malnutrition_zambia$zscore,
  xlab = "child's age in month",
  ylab = "Z-score"
)

boxplot(
  zscore ~ c_age,
  data = malnutrition_zambia,
  xlab = "child's age in month",
  ylab = "Z-score",
  xaxt = "n"
)
```

```{r, fig.width=10, fig.height=12, fig.align='center'}
par(mfrow = c(3, 2))

boxplot(
  zscore ~ c_breastf,
  data = malnutrition_zambia,
  xlab = "duration of breastfeeding in months",
  ylab = "average Z-score",
  xaxt = "n"
)

malnutrition_zambia$bmi_group <- cut(
  malnutrition_zambia$m_bmi,
  breaks = seq(
    min(malnutrition_zambia$m_bmi),
    max(malnutrition_zambia$m_bmi),
    length.out = 101
  ),
  include.lowest = TRUE
)

boxplot(
  zscore ~ bmi_group,
  data = malnutrition_zambia,
  xlab = "mother's BMI",
  ylab = "average Z-score",
  xaxt = "n"
)

malnutrition_zambia$age_group <- cut(
  malnutrition_zambia$m_agebirth,
  breaks = seq(
    min(malnutrition_zambia$m_agebirth),
    max(malnutrition_zambia$m_agebirth),
    length.out = 101
  ),
  include.lowest = TRUE
)

boxplot(
  zscore ~ age_group,
  data = malnutrition_zambia,
  xlab = "mother's age in years",
  ylab = "average Z-score",
  xaxt = "n"
)

malnutrition_zambia$height_group <- cut(
  malnutrition_zambia$m_height,
  breaks = seq(
    min(malnutrition_zambia$m_height),
    max(malnutrition_zambia$m_height),
    length.out = 101
  ),
  include.lowest = TRUE
)

boxplot(
  zscore ~ height_group,
  data = malnutrition_zambia,
  xlab = "mother's height in cm",
  ylab = "average Z-score",
  xaxt = "n"
)

boxplot(
  zscore ~ m_education,
  data = malnutrition_zambia,
  xlab = 'mother\'s level of education',
  ylab = 'Z-score',
  names = c('none', 'primary', 'secondary', 'higher')
)
```
