---
title: "Extensions of the Classical Linear Model"
output: html_notebook
---

## Example 4.1 Munich Rent Index — Diagnostics for Heteroscedastic Errors

Importing the Munich Rent Index data:

```{r}
source("import_data/munich_rent_index.R")
```

Adjust the model:

$$
\text{rent}_{i} = \beta_{0} + \beta_{1} \text{area}_{i} + \beta_{2} \text{yearco}_{i} + \beta_{3} \text{yearco2}_{i} + \beta_{4} \text{yearco3}_{i} + \epsilon_{i}
$$

```{r}
munich_rent_poly_yearc <- poly(munich_rent_index$yearc, 3)
munich_rent_index[, c("yearco", "yearco2", "yearco3")] <- munich_rent_poly_yearc
munich_regression <- lm(rent ~ area + yearco + yearco2 + yearco3, data = munich_rent_index)
summary(munich_regression)
```

And we visualize:


```{r, fig.width=10, fig.height=12, fig.align='center'}
par(mfrow = c(3, 2))

area_seq <- seq(min(munich_rent_index$area), max(munich_rent_index$area), length.out = 100)
yearc_seq <- seq(min(munich_rent_index$yearc), max(munich_rent_index$yearc), length.out = 100)

munich_beta <- munich_regression$coefficients

area_effect <- function(area){
  munich_beta[2] * area
}

yearc_effect <- function(yearc){
  yearc_poly <- predict(munich_rent_poly_yearc, yearc)
  munich_beta[3] * yearc_poly[, 1] + munich_beta[4] * yearc_poly[, 2] + munich_beta[5] * yearc_poly[, 3]
}

partial_area <- munich_rent_index$rent - munich_beta[1] - yearc_effect(munich_rent_index$yearc)
plot(
  munich_rent_index$area,
  partial_area,
  main = 'effect of area incl. partial residuals',
  xlab = 'area in sqm',
  ylab = 'effect of area'
)
lines(area_seq, area_effect(area_seq), col = 'red', lwd = 2)

partial_year <- munich_rent_index$rent - munich_beta[1] - area_effect(munich_rent_index$area)
plot(
  munich_rent_index$yearc,
  partial_year,
  main = 'effect of year of construction incl. partial residuals',
  xlab = 'year of construction',
  ylab = 'effect of year of construction'
)
lines(yearc_seq, yearc_effect(yearc_seq), col = 'red', lwd = 2)

studentized_residuals <- rstudent(munich_regression)
plot(
  munich_regression$fitted.values,
  studentized_residuals,
  main = 'studentized residuals vs. estimated rent',
  xlab = 'estimated rent',
  ylab = 'studentized residuals'
)
plot(
  munich_rent_index$area,
  studentized_residuals,
  main = 'studentized residuals vs. area',
  xlab = 'area in sqm',
  ylab = 'studentized residuals'
)
plot(
  munich_rent_index$yearc,
  studentized_residuals,
  main = 'studentized residuals vs. year of construction',
  xlab = 'year of construction',
  ylab = 'studentized residuals'
)
```

## Example 4.2 Munich Rent Index — Breusch–Pagan Test

Assuming the variance model:

$$
{\sigma_{i}}^{2} = \sigma_{i} \cdot h \left ( \alpha_{0} + \alpha_{1} \text{areao}_{i} + \alpha_{2} \text{areao2}_{i} + \alpha_{3} \text{areao3}_{i} + \alpha_{4} \text{yearco}_{i} + \alpha_{5} \text{yearco2}_{i} + \alpha_{6} \text{yearco3}_{i} \right )
$$

We add the polynomial of degree 3 of $\text{area}$:

```{r}
munich_rent_poly_area <- poly(munich_rent_index$area, 3)
munich_rent_index[, c("areao", "areao2", "areao3")] <- munich_rent_poly_area
```

We calculate the Breusch and Pagan Test:

$$
{\sigma_{i}}^{2} = \sigma_{i} \cdot h \left ( \alpha_{0} + \alpha_{1} {z}_{i1} + \cdots + \alpha_{q} {z}_{iq} \right )
$$

$$
\begin{matrix}
  H_{0}: \alpha_{1} = \cdots = \alpha_{q} = 0 & \text{against} & H_{1}: \alpha_{j} \neq 0 \text{ for at least one } j
\end{matrix}
$$

$$
g_{i} = \frac{{\hat{\epsilon_{i}}}^{2}}{{\hat{\sigma}}_{ML}^{2}}
$$

$$
T = \frac{1}{2} \sum_{i=1}^{2} \left ( \hat{g}_{i} - \bar{g} \right )^{2}
$$

We calculate using R package `lmtest`:

```{r}
library(lmtest)
bptest(
  munich_regression,
  ~ areao + areao2 + areao3 + yearco + yearco2 + yearco3,
  data = munich_rent_index
)
```

## Example 4.3 Munich Rent Index — Two-Stage Estimation

### First step: Classical linear model

Using the model of Example 4.1:

```{r}
summary(munich_regression)
```

### Second step: Auxiliary regression

Estimate the regression model of example

$$
\log({\hat{\epsilon}_{i}}^{2}) = \alpha_{0} + \alpha_{1} \text{areao}_{i} + \alpha_{2} \text{areao2}_{i} + \alpha_{3} \text{areao3}_{i} + \alpha_{4} \text{yearco}_{i} + \alpha_{5} \text{yearco2}_{i} + \alpha_{6} \text{yearco3}_{i} + \nu_{i}
$$

```{r, fig.width=10, fig.height=8, fig.align='center'}
par(mfrow = c(2, 2))

log_sq_residuals <- log(residuals(munich_regression)^2)
variance_model <- lm(
  log_sq_residuals ~ areao + areao2 + areao3 + yearco + yearco2 + yearco3,
  data = munich_rent_index
)
summary(variance_model)

plot(
  munich_rent_index$area,
  log_sq_residuals,
  main = 'log. squared residuals vs. area',
  xlab = 'log. squared residuals',
  ylab = 'area in sqm'
)

plot(
  munich_rent_index$yearc,
  log_sq_residuals,
  main = 'log. squared residuals vs. year of construction',
  xlab = 'log. squared residuals',
  ylab = 'year of construction'
)

variance_beta <- variance_model$coefficients

area_variance_effect <- function(area){
  i_poly_area <- predict(munich_rent_poly_area, area)
  variance_beta[2] * i_poly_area[, 1] + variance_beta[3] * i_poly_area[, 2] + variance_beta[4] * i_poly_area[, 3]
}

yearc_variance_effect <- function(yearc){
  yearc_poly <- predict(munich_rent_poly_yearc, yearc)
  variance_beta[5] * yearc_poly[, 1] + variance_beta[6] * yearc_poly[, 2] + variance_beta[7] * yearc_poly[, 3]
}

plot(
  munich_rent_index$area,
  log_sq_residuals - variance_beta[1] - yearc_variance_effect(munich_rent_index$yearc),
  main = 'residual regression: effect of area',
  xlab = 'effect of area',
  ylab = 'area in sqm'
)
lines(area_seq, area_variance_effect(area_seq), col = 'red', lwd = 2)

plot(
  munich_rent_index$yearc,
  log_sq_residuals - variance_beta[1] - area_variance_effect(munich_rent_index$area),
  main = 'residual regression: effect of year of construction',
  xlab = 'effect of year of construction',
  ylab = 'year of construction'
)
lines(yearc_seq, yearc_variance_effect(yearc_seq), col = 'red', lwd = 2)
```

### Third step: Weighted linear model

We calculate the Weight:

```{r}
weight <- 1 / exp(variance_model$fitted.values)
head(weight)
```

And now the adjustment:

```{r}
weighted_linear_model <- lm(
  rent ~ area + yearco + yearco2 + yearco3,
  data = munich_rent_index,
  weights = weight
)
summary(weighted_linear_model)
```

And we compare the covariance matrix of both model:

```{r}
vcov(munich_regression)
vcov(weighted_linear_model)
```

And the confidence intervals:

```{r}
linear_confint <- confint(munich_regression)
linear_confint
weighted_confint <- confint(weighted_linear_model)
weighted_confint
```

## Example 4.4 Munich Rent Index — White-Estimator

We use R to obtain the covariance matrix and the confidence intervals:

```{r}
library(sandwich)
robust_se <- sqrt(diag(vcovHC(munich_regression, type = "HC1")))
coefs <- munich_regression$coefficients
z <- qnorm(1 - 0.05 / 2)
lower <- coefs - z * robust_se
upper <- coefs + z * robust_se
confint_robust <- data.frame(lower, upper)
colnames(confint_robust) <- colnames(linear_confint)
confint_robust
```

## Example 4.5 Prices of Used Cars — Weighted Regression
